rawPaper

### 2020-11-02 analysis/paper/rawPaper.Rmd
ready to modelTesting-OK

gapsFasta <- read.fasta(file = "../data/raw_data/Sicydium_COI_alignment.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
replacingGaps <- lapply(gapsFasta, function(x) str_replace_all(x, "-", "N"))
write.fasta(replacingGaps,names(replacingGaps),"../data/raw_data/Sicydium_COI_gaps.fasta")

@@

gettingRF <- function(x){
  fasta <- readDNAStringSet("../data/raw_data/Sicydium_COI_gaps.fasta")
  orf1 <- predORF(fasta[1:10], n=1, type="df", mode="orf", strand="sense")   # predORF() do not read seqs with gaps
  # predicts longest ORF for sense strand in each query sequence                                                                          
  longestSeq1 <- orf$start[orf$width == max(orf$width)]
  codonPosition1 <- longestSeq[1]/3
  codonConfirmation1 <- round(codonPosition1)
  if (codonPosition1==codonConfirmation1){
    print("The Reading Frame of the seqs is +1, no further action is required, go to modelTesting")
    } else {
    if (codonPosition1!=codonConfirmation1){
      #remove first nucleotide position
      newRF2 <- subseq(fasta, start = 2, end = tempLength)      # 651 bp

      # getting new RF from the new alignment with -1 bp.
      orf2 <- predORF(newRF2[1:10], n=1, type="df", mode="orf", strand="sense")

      # checking size gain
      longestSeq2 <- orf2$start[orf2$width == max(orf2$width)]
      codonPosition2 <- (longestSeq2[1]+2)/3
      codonConfirmation2 <- round(codonPosition2)
      if (codonPosition2==codonConfirmation2){
        writeXStringSet(newRF2, file="../data/raw_data/Sicydium_COI_newRF2.fasta")
      print("The Reading Frame of the seqs is +2, new fasta file created, go to modelTesting")
      } else {
      if (codonPosition2!=codonConfirmation2){
        #remove first nucleotide position
        tempLength <- length(fasta[[1]])
        newRF3 <- subseq(fasta, start = 3, end = tempLength)
        orf3 <- predORF(newRF3[1:10], n=1, type="df", mode="orf", strand="sense") 
        # checking size gain
        longestSeq3 <- orf3$start[orf3$width == max(orf3$width)]
        codonPosition3 <- (longestSeq3[1]+2)/3
        codonConfirmation3 <- round(codonPosition3)
        if (codonPosition3==codonConfirmation3){
          writeXStringSet(newRF3, file="../data/raw_data/Sicydium_COI_newRF3.fasta")
        print("The Reading Frame of the seqs is +3, new fasta file created, go to modelTesting")
       }
     }
   }  
    }
  }
}

@@

#getting all packages and libraries
#install.packages("modelTest")
#install.packages("ape")
#library(ape)

# reading the file again, only if necessary
var1 <- read.dna(sequences already alignned, format = 'fasta')    # format = can be sequential, interleaved, clustal
                                                                  # as.character = TRUE/FALSE, default is class DNAbin

@@

#getting the class used by modelTest: phyDat or pml
var2 <- phyDat(var1, type = "DNA", levels = NULL)

# getting the best model
var3 <- modelTest(var2, tree = NULL, model = "all", multicore = TRUE, mc.cores = 2)

# getting the AICc for each competing model
# visually inspecting the AICc is better than just getting the best model automatically, because it allows to compares models with similar AICc and it is my personal opinion to choose the one with the least parameters.  
print(var3)

# storing the best model
var4 <- dist.ml(var3, model = "uno d esos")

>>>>

gapsFasta <- read.fasta(file = "../data/raw_data/Sicydium_COI_dupli.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
replacingGaps <- lapply(gapsFasta, function(x) str_replace_all(x, "-", "N"))
write.fasta(replacingGaps,names(replacingGaps),"../data/raw_data/Sicydium_COI_gaps.fasta")

@@

gettingRF <- function(x){
  fasta <- readDNAStringSet("../data/raw_data/Sicydium_COI_gaps.fasta")
  orf1 <- predORF(fasta[1:10], n=1, type="df", mode="orf", strand="sense")   # predORF() do not read seqs with gaps
  # predicts longest ORF for sense strand in each query sequence                                                                          
  longestSeq1 <- orf1$start[orf1$width == max(orf1$width)]
  codonPosition1 <- longestSeq1[1]/3
  codonConfirmation1 <- round(codonPosition1)
  if (codonPosition1==codonConfirmation1){
   file.rename("../data/raw_data/Sicydium_COI_gaps.fasta", "../data/raw_data/Sicydium_COI_newRF.fasta") 
  print("The Reading Frame of these seqs is +1, no further action is required, go to modelTesting")
    } else {
    if (codonPosition1!=codonConfirmation1){
      tempLength <- length(fasta[[1]])
      #remove first nucleotide position
      newRF2 <- subseq(fasta, start = 2, end = tempLength)      # 651 bp
      # getting new RF from the new alignment with -1 bp.
      orf2 <- predORF(newRF2[1:10], n=1, type="df", mode="orf", strand="sense") 
      # checking size gain
      longestSeq2 <- orf2$start[orf2$width == max(orf2$width)]
      codonPosition2 <- (longestSeq2[1]+2)/3
      codonConfirmation2 <- round(codonPosition2)
      if (codonPosition2==codonConfirmation2){
        writeXStringSet(newRF2, file="../data/raw_data/Sicydium_COI_newRF.fasta")
      print("The Reading Frame of these seqs was +2, new fasta file created, go to modelTesting")
      } else {
      if (codonPosition2!=codonConfirmation2){
        #remove first nucleotide position
        tempLength <- length(fasta[[1]])
        newRF3 <- subseq(fasta, start = 3, end = tempLength)
        writeXStringSet(newRF3, file="../data/raw_data/Sicydium_COI_newRF.fasta")
      print("The Reading Frame of these seqs was +3, new fasta file created, go to modelTesting")
     }
   }  
  }
 }
}

@@

#getting all packages and libraries
BiocManager::install("modelTest")
install.packages("ape")
#library(modelTest)

install.packages("phangorn")
library(phangorn)

gettingBestModel <- function(x){
  var1 <- read.dna(file="../data/raw_data/Sicydium_COI_newRF.fasta", format = "fasta")
  var2 <- as.phyDat(var1)
  var3 <- modelTest(var2, tree = NULL, model = c("JC", "K80", "HKY", "TrN", "SYM", "GTR"), multicore = TRUE, mc.cores = 2)
  return(var3)
}

gettingBestModel(x)

mammals <- read.dna("mammals.dna", format="interleaved")
mammals_phyDat <- phyDat(mammals, type = "DNA", levels = NULL)
mammals10 <- subset(mammals_phyDat, 1:10)
mammals10_phyDat <- phyDat(mammals10, type = "DNA", levels = NULL)

var1 <- read.dna(file="../data/raw_data/Sicydium_COI_newRF.fasta", format = "fasta")
var2 <- as.phyDat(var1)

var11 <- subset(var2, 1:10)
var21 <- phyDat(var11, type = "DNA", levels = NULL)

var3 <- modelTest(var21, tree = NULL, model = c("JC", "K80", "HKY", "TrN", "SYM", "GTR"), multicore = TRUE, mc.cores = 2)

print(var3)

dna_dist <- dist.ml(var11, model="JC69")
mammals_NJ  <- NJ(dna_dist)
plot(mammals_NJ, main = "Neighbor Joining")

fit <- pml(mammals_NJ, var11)
print(fit)

fitJC <- optim.pml(fit, model = "JC", rearrangement = "stochastic")
bs <- bootstrap.pml(fitJC, bs=100, optNni=TRUE, multicore=TRUE, control = pml.control(trace=0))
plotBS(midpoint(fitJC$tree), bs, p = 50, type="p")
write.tree(bs, file="bootstrap_example.tre")

# format = can be sequential, interleaved, clustal
# as.character = TRUE/FALSE, default is class DNAbin

@@

#getting the class used by modelTest: phyDat or pml
# getting the best model
# getting the AICc for each competing model
# visually inspecting the AICc is better than just getting the best model automatically, because it allows to compares models with similar AICc and it is my personal opinion to choose the one with the least parameters.
